---
title: "Brownlows Bets"
author: "Tracy Alset"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
source(file="BrLw.R")
options( digits=2)
```


<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document
 
Report git commit structure:
There are 3 questions and there should be at least 3 separate commits, at least one dealing with each of the 3 questions.
1. What are the fair odds for a bet on matching birthdays in a 35 person class.
2. For what size class is a 2:1 odds bet fair.
3. How much money would Brownlow be expected to end up with if the bets went as described.

## Our strategy is to build a Function and the use it to answer the three questions. 

The function will be given a class_size by number of students, odds, and Number_of_times brownlow made the bet.  It will give you back how much Brownlows earnings were. We will use it to answer our three questions below. If you can, improve this function by giving default values to some of its inputs. 

# In the class and payoff as described  If you highered 40 brownlow substitute teachers to do the same thing that Brownlow did then.

```{r}
Substitute<- numeric(length=40)
for (j in 1:40) {
Substitute[j]=BrLw(odds=2)  
}

```


## what is their average earnings

```{r}
mean(Substitute)
```

The average earnings is `r mean(Substitute) ` from having 40 brownlows do the same thing.

## the standard deviation and a 95% confidence interval for their earnings?
  
<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows winnings (or losings)-->


```{r}
sd(Substitute)
```
The standard deviation is `r sd(Substitute) ` and based on the central limit theorem by taking two standard deviations from the left ( from the mean)and adding two standard deviations to the mean , that's where the 95% confidence interval is which would be from `r mean(Substitute)- 2*sd(Substitute)` to `r mean(Substitute) + 2*sd(Substitute)` 

# What are the fair odds? Use manipulation of odds in the function to tune Brownlows earnings to zero. 

```{r}
Fair<- numeric(length=1000)
for (j in 1:1000) {
Fair[j]=BrLw(odds=4.38)  
}

```


```{r}
hist(Fair)
```


Looking at the histogram we see that it is centered around zero showing that the profit is zero ( having fair odds), and that happens when brownlow wagers $4.38 for every $1 that the students wager. 


<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 637 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->




# What size of class makes the 2:1 bet fair?  Use manipulation of number of Students to tune Brownlows earnings to zero.


```{r}
Sub<- numeric(length=1000)
for (j in 1:1000) {
Sub[j]=BrLw(numClass = 28 )  
}

```

```{r}
hist(Sub)
```

It looks like when we have 28 students in the class the 2:1 odd is fair because the peak is close to zero. To be more precise, the actual number that would make the histogram center at zero would be between 28 and 29, but because we're talking about people we should have an integer for the number of students in the class. (In the histogram , the zero was further to the right in the case of 29 students and this shift is larger when we use 28 as the size of the class where the zero was on the left but the shift is smaller than the case of having 29 as the class size).


<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting profits!-->


# Discussion and Conclusion

The first thing that we did was create a function that allows us to find brownlow's earnings based on the class size, the odds and the amount of times brownlow does the bet. By manipulating this function we were able to answer the questions that we had to answer. first we were able to find the earnings of hiring 40 brownlows to do the same thing by finding the mean (after using the function we created )and then the standard deviation to find the 95% confidence interval for the earnings. then we were able to find the fair odds( by changing the odds in the function) by manipulating the function. and then the last thing we did was manipulate the function ( by changing the class size) to find where the earnings are zero in case of a 2:1 odds.



<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->

